---
title: "RFM Segmentation"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r init}
library(tidyverse)
library(gridExtra)

rfm <- read_csv("../cleaned_data/rfm_2020-03-29_to_2020-06-14.csv", col_types="_ciiii")
demographics <- readxl::read_excel("../data/DSA3101_Hackathon_Panelists_Demographics.xlsx")
head(rfm)
head(demographics)
```

```{r viz}
full_info <- rfm %>% left_join(demographics, by=c("Panel ID"="ID"))
full_info %>% group_by(R, F, M) %>% summarise(count=n()) %>% 
  ggplot() + geom_point(aes(x=R, y=F, size=count)) + facet_wrap("M")

```

```{r}
full_info %>% filter(Income=="Income >5000") %>% group_by(R, F, M) %>% summarise(count=n()) %>% ggplot() + geom_point(aes(x=R, y=F, size=count)) + facet_wrap("M")
```

```{r}
high_income <-full_info %>% filter(Income == "Income >5000")
```

## Examining High-Income group

From the 2 plots below it shows that the higher-income group is mainly made up of those from Central M'sia, and predominantly Chinese.

```{r}
full_info %>% ggplot(mapping=aes(x=Ethnicity)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))
full_info %>% filter(Income == "Income >5000") %>% ggplot(mapping=aes(x=Ethnicity)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))

```

## Examining subgroup - High-Income, but low spending

We define this group to be those that have above average R/F scores but have low M value. Ideally we would be able to understand why this group with high-purchasing power spends little with us, and how/why they differ from the 333-444s.

```{r}
high_income_low_spenders <- full_info %>% 
  filter(Income== "Income >5000", R>2, F>2, M<3)

high_income_low_spenders %>% ggplot(mapping=aes(x=Ethnicity)) + 
  geom_bar() + theme(axis.text.x=element_text(angle=90))
```

No significant difference in ethnic composition

```{r}
p1 <- high_income %>% ggplot(mapping=aes(x=Lifestage)) + 
  geom_bar() + ggtitle("High Spenders") + theme(axis.text.x=element_text(angle=90))
p2 <- high_income_low_spenders %>% ggplot(mapping=aes(x=Lifestage)) + geom_bar() + ggtitle("Lower Spenders") + theme(axis.text.x=element_text(angle=90))

grid.arrange(p1, p2, nrow=1)
```

Also does not seem to be a particular lifestage group that accounts for this group of spenders.

```{r}
p1 <- high_income %>% ggplot(mapping=aes(x=BMI)) + 
  geom_bar() + ggtitle("High Spenders") + theme(axis.text.x=element_text(angle=90))
p2 <- high_income_low_spenders %>% ggplot(mapping=aes(x=BMI)) + geom_bar() + ggtitle("Lower Spenders") + theme(axis.text.x=element_text(angle=90))

grid.arrange(p1, p2, nrow=1)
```


```{r}
p1 <- high_income %>% ggplot(mapping=aes(x=`#HH`)) + 
  geom_bar() + ggtitle("High Spenders") + theme(axis.text.x=element_text(angle=90))
p2 <- high_income_low_spenders %>% ggplot(mapping=aes(x=`#HH`)) + geom_bar() + ggtitle("Lower Spenders") + theme(axis.text.x=element_text(angle=90))

grid.arrange(p1, p2, nrow=1)
```
